---
title: "preprocessing_hcc"
output: html_document
date: "2024-01-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#0.Preparation
```{r}
library(RnBeads)
dir <-  getwd()
dataDir <- file.path(dir, "data")
idatDir <- file.path(dataDir, "idat")
sampleSheet <- file.path(dataDir, "samplesheet_final.csv")
dataSource <- c(idatDir, sampleSheet)
resultDir <- file.path(dir, "results")
reportDir <- file.path(resultDir, "report")
imageDir <- file.path(dir, "preprocessing_hcc.RData")
save.rnb.set(rnb, path = dir, archive = FALSE)
gc()
```

#0. COmputing management
```{r}
parallel.isEnabled()
parallel.setup(2)
parallel.isEnabled()
#parallel.terdown()
```



#1.Import
```{r}
rnb <- rnb.execute.import(dataSource, data.type="infinium.idat.dir")
save.image(imageDir)

```
#2. Filter NA
```{r}

# Identify rows with at least one NA value
rows_with_na <- apply(Mval.filter, 1, function(row) any(is.na(row)))

# Calculate the percentage of NA values in each row with at least one NA
percent_na <- apply(Mval.filter[rows_with_na, ], 1, function(row) {
  na_count <- sum(is.na(row))
  total_columns <- length(row)
  percent_na <- (na_count / total_columns) * 100
  return(percent_na)
})

# Print the result
print(percent_na)
```



#2.Filter
```{r}
rnb.filtering <- rnb
cat("Initial Probes: " ,nrow(meth(rnb.filtering))) # the number of sites in the unfiltered object
cat("Initial Sample: " ,ncol(meth(rnb.filtering)))
#.p-Value (optional)
#any.bad.p.val <- apply(dpval(rnb.set.unfiltered)>0.01, 1, any)
#rnb.set.unfiltered <- remove.sites(rnb.set.unfiltered, any.bad.p.val)
#nsites(rnb.set.unfiltered)

# 2.1 Remove probes outside of CpG context
rnb.filterCpG <- rnb.execute.context.removal(rnb.filtering)
rnb.filtering <- rnb.filterCpG$dataset
cat("CpG Probes: " ,nrow(meth(rnb.filtering))) # the number of CpG sites in the unfiltered object

# 2.2 SNP filtering allowing no SNPs in the probe sequence
rnb.filterSNP<- rnb.execute.snp.removal(rnb.filtering, snp="any")
rnb.filtering <- rnb.filterSNP$dataset
cat("nonSNP Probes: " ,nrow(meth(rnb.filtering))) 

# 2.3 Remove CpGs on sex chromosomes
rnb.filterSex <- rnb.execute.sex.removal(rnb.filtering)
rnb.filtering <- rnb.filterSex$dataset
cat("nonSex Chromosome: " ,nrow(meth(rnb.filtering))) 

# 2.4 Remove probes and samples based on a greedy approach
greedycut <- rnb.execute.greedycut(rnb.filtering)
rnb.filtering <- remove.samples(rnb.filtering, greedycut$samples) #filter samples
rnb.filtering <- remove.sites(rnb.filtering, greedycut$sites)
cat("greedycut probes: " ,nrow(meth(rnb.filtering))) 

# 2.5 Remove probes containing NA for beta values
rnb.filterNA <- rnb.execute.na.removal(rnb.filtering, threshold = 0.1)
rnb.filtering <- rnb.filterNA$dataset
cat("nonNA Probes: " ,nrow(meth(rnb.filtering))) 

# 2.6 Remove probes for which the beta values have low standard deviation
rnb.filterSTD <- rnb.execute.variability.removal(rnb.filtering, 0.005)
rnb.filtering <- rnb.filterSTD$dataset
cat("nonVar Probes: " ,nrow(meth(rnb.filtering))) 

# 2.7 Imputation
#rnb.filtering <- rnb.execute.imputation(rnb.filtering,
#                                           method = "knn",
#                                           update.ff = TRUE)
cat("Final Probes Count: " ,nrow(meth(rnb.filtering)))
cat("Final Sample Count: " ,ncol(meth(rnb.filtering))) 
save.image(imageDir2)
```
rnb.filtering <- rnb
> cat("Initial Probes: " ,nrow(meth(rnb.filtering))) # the number of sites in the unfiltered object
Initial Probes:  485577
> cat("Initial Sample: " ,ncol(meth(rnb.filtering)))
Initial Sample:  452
> #.p-Value (optional)
> #any.bad.p.val <- apply(dpval(rnb.set.unfiltered)>0.01, 1, any)
> #rnb.set.unfiltered <- remove.sites(rnb.set.unfiltered, any.bad.p.val)
> #nsites(rnb.set.unfiltered)
> 
> # 2.1 Remove probes outside of CpG context
> rnb.filtering <- rnb.execute.context.removal(rnb.filtering)$dataset
> cat("CpG Probes: " ,nrow(meth(rnb.filtering))) # the number of CpG sites in the unfiltered object
CpG Probes:  482421
> 
> # 2.2 SNP filtering allowing no SNPs in the probe sequence
> rnb.filtering <- rnb.execute.snp.removal(rnb.filtering, snp="any")$dataset
> cat("nonSNP Probes: " ,nrow(meth(rnb.filtering))) 
nonSNP Probes:  403331
> 
> # 2.3 Remove CpGs on sex chromosomes
> rnb.filtering<- rnb.execute.sex.removal(rnb.filtering)$dataset
> cat("nonSex Chromosome: " ,nrow(meth(rnb.filtering))) 
nonSex Chromosome:  392643
> 
> # 2.4 Remove probes and samples based on a greedy approach
> greedycut <- rnb.execute.greedycut(rnb.filtering)
Warning: 'memory.size()' is no longer supported2024-01-08 10:09:02     Inf    INFO Working with a p-value threshold of 0.05
Warning: 'memory.size()' is no longer supported2024-01-08 10:11:48     Inf  STATUS Calculated a total of 15242 iterations
Warning: 'memory.size()' is no longer supported2024-01-08 10:11:48     Inf    INFO Optimal number of iterations is 15242
> rnb.filtering <- remove.samples(rnb.filtering, greedycut$samples, verbose = TRUE) #filter samples
Error in remove.samples(rnb.filtering, greedycut$samples, verbose = TRUE) : 
  unused argument (verbose = TRUE)
> cadangan <- rnb.filtering
> rnb.filtering <- remove.samples(rnb.filtering, greedycut$samples) #filter samples
> rnb.filtering <- remove.sites(rnb.filtering, greedycut$sites)
> cat("greedycut probes: " ,nrow(meth(rnb.filtering))) 
greedycut probes:  377407
> 
> # 2.5 Remove probes containing NA for beta values
> rnb.filtering<- rnb.execute.na.removal(rnb.filtering, 
+                                            threshold = 0.1)$dataset
> cat("nonNA Probes: " ,nrow(meth(rnb.filtering))) 
nonNA Probes:  377403
> 
> # 2.6 Remove probes for which the beta values have low standard deviation
> rnb.filtering <- rnb.execute.variability.removal(rnb.filtering, 0.005)$dataset
> cat("nonVar Probes: " ,nrow(meth(rnb.filtering))) 
nonVar Probes:  377403
> 
> # 2.7 Imputation
> rnb.filtering <- rnb.execute.imputation(rnb.filtering,
+                                            method = "knn",
+                                            update.ff = TRUE)
Warning: 'memory.size()' is no longer supported2024-01-08 12:27:59     Inf  STATUS STARTED Imputation procedure knn 
Warning: 'memory.size()' is no longer supported2024-01-08 13:13:05     Inf  STATUS COMPLETED Imputation procedure knn 

> cat("Final Probes Count: " ,nrow(meth(rnb.filtering)))
Final Probes Count:  377403
> cat("Final Sample Count: " ,ncol(meth(rnb.filtering))) 
Final Sample Count:  447
> save.image(imageDir)


#3.Normalization
```{r}
rnb.norm <- rnb.filtering
rnb.norm <- rnb.execute.normalization(rnb.norm, method="bmiq",
                                      bgcorr.method="methylumi.noob")
save.image(imageDir2)
```

#4.imputation
```{r}
# 2.7 Imputation
rnb.norm.impute <- rnb.execute.imputation(rnb.norm, method = "knn", update.ff = TRUE)
```
#5. SVA
```{r}
svaObj <- rnb.execute.sva(rnb.norm.impute, cmp.cols="Sample_Group", numSVmethod="be")
rnbs.sva <- set.covariates.sva(rnb.norm.impute, svaObj)
get.covariates.sva(rnbs.sva, "Sample_Group")
```

#6. DMP
```{r}
dm.sva <- rnb.execute.computeDiffMeth(rnbs.sva, "Sample_Group")
```

